### 2023-01-16 paper.tex
add review and background on kernel ridgeless estimators

\documentclass[12pt]{amsart}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{hyperref}
\usepackage[nopatch=eqnum]{microtype}
\usepackage{cleveref}

\newcommand{\R}{\mathbb{R}}

\title{Wiggling the bandwidth in ridgeless regression 
    with scale-dependent kernels}
\author{Luca Wellmeier}
\begin{document}
\begin{abstract}
    We explore the possibility of using two different bandwidths in the 
    fitting and evaluation part in kernel ridgeless regression.
    The provided experimental results indicate that \textit{wiggling}
    the bandwidth in this way can generalize and be a cheap alternative to
    the classical Tikhonov square-norm penality.
    Finally, we propose and test a new iterative method for optimizing the 
    bandwidth parameter by employing wiggling in a local search.
\end{abstract}
\maketitle
\tableofcontents

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Let $\mathbf X \in \R^{n \times d}$ be a matrix of $n$ training examples
$X_1, \dots, X_n \in \R^d$ in $d$-dimensional Euclidean space as rows and 
let $Y$ be the column vector of responses $y_1, \dots, y_n \in \R$.
The main actor of this paper is the kernel ridgeless regression estimator
that predicts the response on an unseen data point $X \in \R^d$ as
\begin{equation} \label{eq:ridgeless}
    \hat f(X) = \sum_{i=1}^n (\mathbf K^\dagger Y)_i K(X, X_i).
\end{equation}
Here, $mathbf K$ denotes the kernel matrix $\mathbf K_{ij} = k(X_i, X_j)$
is the \emph{kernel matrix} obtained by evaluating a kernel function 
$k \colon \R^d \times \R^d \to \R$ on the training examples and 
$\mathbf K^\dagger$ denotes its \emph{pseudo-inverse}.
We briefly recall the motivations behind these notions.
First, the \emph{representer theorem} asserts that any minimizer to the 
least squares problem
\[ \min_{f \in \mathcal H_k} \frac 1n \sum_{i=1}^n (y_i - f(X_i))^2 \]
in the \emph{reproducing kernel Hilbert space} $\mathcal H_k$ associated
to $k$, is given by a linear combination of the form
$\sum_{i=1}^n c_i k(X,X_i)$.
This allows to apply the usual linear least squares method to a - at the 
first glance - non-linear approximation problem in a possibly 
infinite-dimensional hypothesis space by pushing the data points into 
the \emph{feature space} spanned by $K(X, \cdot) \in \mathcal H_k$ 
for $X \in \R^d$.
The kernel matrix in this setting corresponds to the empirical covariance
matrix $\mathbf X^T \mathbf X$ and classical theory suggests that, even if 
it does not possess full-rank, there will always exist a minimizer to 
the least squares problem above.
The pseudo-inverse then corresponds to the unique least-squares solution
with minimal $\mathcal H_k$-norm.
This reduction of an infinite-dimensional optimization problem to a
finite-dimensional one using kernels is generally known as the 
\emph{kernel trick}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Wiggling as a regularization technique}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Iterative wiggling to find the best bandwidth}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions and directions}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}